PineAPPL:\\ combining EW and QCD corrections for PDF fits

### 2020-08-13 paper/sec-introduction.tex
section 1

The two respects are intertwined: PDFs provide a way of obtaining a prediction
for a given process in terms of other processes, therefore the accuracy and 
precision with which PDFs are able to make predictions crucially depend on 
the accuracy of the perturbative computation used in the PDF determination
(matched to the precision of the data). Electroweak (including combined QCD--EW) 
effects should therefore naturally be included in a PDF determination. The 
reason is twofold. First, one expects NNLO QCD and NLO EW corrections to be of 
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in PDF determinations, NLO EW corrections should be
taken into account on the same footing. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. Such a behaviour becomes relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

The systematic inclusion of EW corrections in a PDF determination entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are a fundamental ingredient
in any PDF fit, where the evaluation of the hadronic cross section needs to 
be performed a large number of times. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate processes beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a new computer library that provides a direct interface
from parton-level Monte Carlo event generators accurate up to any fixed order to
fast-interpolation grids. The library supports variations of the factorisation
and renormalisation scales. Planned extensions will include support for resummation and matching with a photon- and/or
parton-shower. An interface to the library is included in
\textsc{MadGraph5\_aMC@NLO} (\textsc{MG5\_aMC} henceforth), with which it has
been developed and tested, and which supersedes \textsc{APPLgrid}~\cite{Carli:2010rw} and \textsc{aMCfast}~\cite{Bertone:2014zva}.
However, \textsc{PineAPPL} can be easily interfaced with every
Monte Carlo generator, e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it could complement \textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

>>>>

The two respects are intertwined: PDFs provide a way of obtaining a prediction
for a given process in terms of other processes, therefore the accuracy and 
precision with which PDFs are able to make predictions crucially depend on 
the accuracy of the perturbative computation used in the PDF determination
(matched to the precision of the data). Electroweak (including combined QCD--EW) 
effects should therefore naturally be included in a PDF determination. The 
reason is twofold. First, one expects NNLO QCD and NLO EW corrections to be of 
comparable size because, at the EW scale, the QCD and EW running couplings 
become similar, $\alphas^2\sim \alpha$. If NNLO QCD corrections are included
by default in PDF determinations, NLO EW corrections should be
taken into account as well. Second, the virtual exchange of soft or 
collinear weak bosons leads to Sudakov 
logarithms~\cite{Denner:2000jv,Denner:2001gw},
which can make the coefficients of the EW series grow faster than 
their QCD counterparts. Such a behaviour becomes relevant in
phase-space regions associated with large mass scales (roughly of the order
of the \si{\tera\electronvolt}), where several LHC data sets (e.g.\ Z-boson
transverse-momentum distributions) enter both the determination of PDFs and the search
for new physics.

The systematic inclusion of EW corrections in a PDF determination entails
the solution of two separate problems. First, a problem of efficiency:
fast-interpolation grids should be constructed, whereby partonic matrix 
elements, accurate to NLO QCD+EW, are precomputed in such a way that the 
numerical convolution with generic input PDFs can be efficiently approximated
by means of interpolation techniques. Such grids are a fundamental ingredient
in any PDF fit, where the evaluation of the hadronic cross section needs to 
be performed a large number of times. While two default formats exist for
these grids, \textsc{APPLgrid}~\cite{Carli:2010rw} and
\textsc{fastNLO}~\cite{Kluge:2006xs,Wobisch:2011ij,Britzger:2012bs}, none of them supports the inclusion of EW
corrections nor the interface to a Monte Carlo generator accurate to 
NLO QCD+EW\@. Second, a problem of consistency: the way in which EW effects may
(or may not) be folded into the data varies across different experimental 
analyses, a fact that challenges their consistent theoretical interpretation. 
Examples are the subtraction of background processes which should not be 
considered as such (e.g.\ the $t$-channel photon-induced component in
neutral-current Drell--Yan, which is not a separate processes beyond leading
order) or of just a part of the EW effects (e.g.\ multiple-photon
radiation from light particles in the final state of neutral- or charged-current
Drell--Yan, especially with electrons). Be that as it may, if EW effects are
systematically included in theoretical predictions, they should not be subtracted
from experimental results, otherwise they will be double counted.

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows the user to generate
fast-interpolation grids, accurate to up to any fixed order in the strong and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed-orders in the strong coupling only) that can be generated 
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{MG5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
\textsc{PineAPPL} can also be easily interfaced with any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it complements
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.
